---
title: "Sensitivity Analysis"
author: "Bryan Clark"
date: "4/1/2021"
output: 
  html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float: true
---

# Goal

Determine which variables the financial model is most sensitive to as it relates to the deal IRR, investor IRR, and sponsor IRR.

# Methodology

One variable at a time, perform a Monte Carlo simulation by adding 5% relative noise. We can then rank/index the variables based on how wide the 95% confidence interval is to determine the most risky assumptions.

# Analysis

## Setup

### R

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
reticulate::use_condaenv("deal-analyzer", required = T)
```

```{r, message=FALSE}
library(googlesheets4)
library(reticulate)
library(tidyverse)
# for exporting data
config = config::get(file = "secrets.yml")
```

### Python

```{python}
import numpy as np
# also here: https://github.com/bclark86/deal-analyzer
from deal_analyzer import generate_deal_specs, DealAnalyzer

deal_specs = generate_deal_specs()
analyzer = DealAnalyzer(**deal_specs)
```

## Baseline

### Deal Specs

```{r}
py$deal_specs
```
### Financials

```{r}
py$analyzer$investor_waterfall
```

## Monte Carlo Simulation 

```{python}
ignore_cols = [
    'building_sf',
    'term',
    'amortization_period',
    'number_of_units',
    'years_held',
    'income_tax_rate'
]
sensitivity_cols = [col for col in deal_specs.keys() if col not in ignore_cols]

np.random.seed(90)
sensitivty_results_df = analyzer.run_sensitivity_analysis(sensitivity_cols, deal_specs)
```

## Summarize Results

```{r}
sensitvity_summary_tbl = py$sensitivty_results_df %>%
    group_by(variable) %>%
    summarize(
        trials = n(),
        lower_95 = quantile(value, 0.025), 
        upper_95 = quantile(value, 0.975),
        deal_irr_lower95 = quantile(deal_irr, 0.025),
        deal_irr_upper95 = quantile(deal_irr, 0.975),
        sponsor_irr_lower95 = quantile(sponsor_irr, 0.025),
        sponsor_irr_upper95 = quantile(sponsor_irr, 0.975),
        investor_irr_lower95 = quantile(investor_irr, 0.025),
        investor_irr_upper95 = quantile(investor_irr, 0.975)
    ) %>%
    mutate(
        deal_irr_ci = deal_irr_upper95 - deal_irr_lower95,
        sponsor_irr_ci = sponsor_irr_upper95 - sponsor_irr_lower95,
        investor_irr_ci = investor_irr_upper95 - investor_irr_lower95
    ) %>%
    mutate(
        deal_risk_index = (scales::rescale(deal_irr_ci, c(0, 1)) * 100) %>% round(0),
        sponsor_risk_index = (scales::rescale(sponsor_irr_ci, c(0, 1)) * 100) %>% round(0),
        investor_risk_index = (scales::rescale(investor_irr_ci, c(0, 1)) * 100) %>% round(0)
    ) %>%
    select(variable, contains("index"), everything())
```

## Export Data

```{r}
gs4_auth(config$stern_email)

sensitvity_summary_tbl %>%
    sheet_write(
        ss = config$sheet_id,
        sheet = "Sensitivity Analysis - Part 1"
    )
```

