---
title: "Sensitivity Analysis v2"
author: "Bryan Clark"
date: "2021-04-10"
output: 
  html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float: true
---

# Goal

Quantify risk of deal in terms of:

1. Investor IRR
2. Covering Investor Preferred Return
3. Cash Flow Positive

# Methodology

Perform a Monte Carlo simulation by sampling values for rent and occupancy rates as they are the most risky assumptions according to the [exploratory sensitivity analysis](https://htmlpreview.github.io/?https://github.com/bclark86/deal-analyzer/blob/main/sensitivity_analysis.html). 

# Analysis

## Setup

### R

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.width = 6,
  fig.asp = 0.618, 
  out.width = "70%", 
  fig.align = "center"
)
reticulate::use_condaenv("deal-analyzer", required = T)
```

```{r, message=FALSE}
library(tictoc)
library(googlesheets4)
library(reticulate)
library(patchwork)
library(tidyverse)
# for exporting data
secrets = config::get(file = "secrets.yml")
theme_set(theme_bw(base_family = "Avenir"))
```

### Python

```{python}
import numpy as np
import pandas as pd
# also here: https://github.com/bclark86/deal-analyzer
from deal_analyzer import (
  generate_deal_specs, 
  DealAnalyzer
)

deal_specs = generate_deal_specs()
analyzer = DealAnalyzer(**deal_specs)
```

## Baseline

### Deal Specs

```{r}
py$deal_specs
```

### Financials

```{r}
py$analyzer$investor_waterfall
```

## Monte Carlo Simulation 

### Sampling Function

```{python}
def sensitivity_analysis_trial(opccupancy_rate_triangular=False):
    # copy the deal specs
    deal_specs_rv = deal_specs.copy()
    # sample random values
    # riskiest inputs are occupancy and rent
    deal_specs_rv['school_year_occupancy_rate'] = (
        np.where(
            opccupancy_rate_triangular,
            np.random.triangular(left=0.86, mode=0.95, right=0.97),
            np.random.uniform(low=0.86, high=0.97)
        ).item()
    )
    deal_specs_rv['summer_occupancy_rate'] = (
        np.where(
            opccupancy_rate_triangular,
            np.random.triangular(left=0.30, mode=0.40, right=0.50),
            np.random.uniform(low=0.30, high=0.5)
        ).item()
    )
    deal_specs_rv['monthly_school_rent'] = (
        np.random.triangular(left=625, mode=650, right=700)
    )
    deal_specs_rv['daily_summer_rent'] = (
        np.random.triangular(left=85, mode=100, right=115)
    )
    # run analyzer
    analyzer_rv = DealAnalyzer(**deal_specs_rv)
    investor_waterfall = analyzer_rv.investor_waterfall
    cfaf_yr1 = investor_waterfall["cash_flow_after_financing"][0]
    investor_preferred_return = investor_waterfall['investor']['preferred_return']
    sponsor_preferred_return = investor_waterfall['sponsor']['preferred_return']
    return {
        # sampled values
        "school_year_occupancy_rate": deal_specs_rv['school_year_occupancy_rate'],
        "summer_occupancy_rate": deal_specs_rv['summer_occupancy_rate'],
        "monthly_school_rent": deal_specs_rv['monthly_school_rent'],
        "daily_summer_rent": deal_specs_rv['daily_summer_rent'],
        # does the deal go into the negative?
        "cfaf_yr1": cfaf_yr1,
        "cash_flow_positive": cfaf_yr1 > 0,
        # can the preferred return be made?
        "investor_preferred_return_covered": cfaf_yr1 > investor_preferred_return,
        "sponsor_preferred_return_covered": cfaf_yr1 > (
            investor_preferred_return + sponsor_preferred_return
        ),
        # investment outcomes
        "deal_irr": investor_waterfall["deal_irr"],
        "investor_irr": investor_waterfall["investor"]["irr"],
        "investor_cash_in": investor_waterfall["investor"]["contribution"],
        "investor_cash_out": investor_waterfall["investor"]["cash_flows"].sum(),
        "sponsor_irr": investor_waterfall["sponsor"]["irr"],
        "sponsor_cash_in": investor_waterfall["sponsor"]["contribution"],
        "sponsor_cash_out": investor_waterfall["sponsor"]["cash_flows"].sum(),
    }
    

def run_simulation(n=100_000, triangular=False):
    np.random.seed(90)
    return pd.DataFrame([
        sensitivity_analysis_trial(triangular) for _ in range(n)
    ])

```

### Run Simulation

```{r}
tic()
results_tbl = tibble(occupancy_distribution = c("triangular", "uniform")) %>%
  mutate(
    function_logical = ifelse(occupancy_distribution == 'triangular', T, F)
  ) %>%
  mutate(
    simulation_data = map(function_logical, .f = ~ {
        py$run_simulation(triangular = .x) %>%
          as_tibble()
      })
  ) %>%
  select(-function_logical)
toc()
```


### Summarize Results

```{r}
summarize_simulation = function(data) {
  data %>%
    as_tibble() %>%
    mutate(
      investor_irr_outcome = case_when(
        investor_irr < 0.10 ~ "less than 10%",
        investor_irr >= 0.10 & investor_irr < 0.15 ~ "10% to 15%",
        investor_irr >= 0.15 & investor_irr < 0.20 ~ "15% to 20%",
        investor_irr >= 0.20 ~ "20% or more"
      ) %>% factor(levels = c("less than 10%", "10% to 15%", "15% to 20%", "20% or more"))
    ) %>%
    group_by(investor_irr_outcome) %>%
    summarize(
      n = n(),
      investor_preferred_return_pct = mean(investor_preferred_return_covered),
      cash_flow_positiv_pct = mean(cash_flow_positive),
      sponsor_irr = median(sponsor_irr)
    ) %>%
    mutate(pct = n / sum(n)) %>%
    relocate(pct, .after = n) %>%
    select(-n, everything())
}

results_tbl$summary = map(results_tbl$simulation_data, summarize_simulation)
```

```{r}
# confidence interval
# probability of preferred return
# bucket summary
investor_irr_ci_tbl = results_tbl %>%
  select(occupancy_distribution, simulation_data) %>%
  unnest(simulation_data) %>%
  group_by(occupancy_distribution) %>%
  summarize(
    investor_irr_05 = quantile(investor_irr, probs = 0.05),
    investor_irr_50 = median(investor_irr),
    investor_irr_95 = quantile(investor_irr, probs = 0.95),
    investor_preferred_return_pct = mean(investor_preferred_return_covered)
  )
```


### Plot Results

```{r}
plot_input_distributions = function(distribution) {
  
  sub_results_tbl = results_tbl %>%
    filter(occupancy_distribution == distribution) %>%
    pluck('simulation_data', 1)
  
  school_year_occupancy_rate = sub_results_tbl %>%
    ggplot(aes(school_year_occupancy_rate)) +
    geom_histogram(binwidth = 0.001) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(
      title = "School Year Occupancy Rate",
      x = "",
      y = ""
    ) + 
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  
  summer_occupancy_rate = sub_results_tbl %>%
    ggplot(aes(summer_occupancy_rate)) +
    geom_histogram(binwidth = 0.001) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(
      title = "Summer Occupancy Rate",
      x = "",
      y = ""
    ) + 
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  
  monthly_school_rent = sub_results_tbl %>%
    ggplot(aes(monthly_school_rent)) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(labels = scales::dollar_format(accuracy = 1)) +
    labs(
      title = "Monthly School Rent",
      x = "",
      y = ""
    ) + 
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  
  daily_summer_rent = sub_results_tbl %>%
    ggplot(aes(daily_summer_rent)) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(labels = scales::dollar_format(accuracy = 1)) +
    labs(
      title = "Daily Summer Rent",
      x = "",
      y = ""
    ) + 
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  
  plots = list(
    school_year_occupancy_rate, summer_occupancy_rate, 
    monthly_school_rent, daily_summer_rent
  )
  
  p = wrap_plots(plots) +
    plot_annotation(title = "Sampling Distributions")
  
  p %>% 
    ggsave(
      filename = str_glue("plots/sampling_distribution_{distribution}.png"),
      width = 8, height = 7
    )
  
  return(p)
}

plot_input_distributions('triangular')
plot_input_distributions('uniform')
```

```{r}
# uniform only for now
simulation_data = results_tbl %>%
  filter(occupancy_distribution == "uniform") %>%
  pluck("simulation_data", 1)

ivesntor_summary_tbl = results_tbl %>%
  filter(occupancy_distribution == "uniform") %>%
  pluck("summary", 1)

investor_irr_ci = investor_irr_ci_tbl %>%
  filter(occupancy_distribution == "uniform")

investor_irr_dist = simulation_data %>%
  mutate(suboptimal = ifelse(investor_irr < 0.10, "yes", "no")) %>%
  ggplot(aes(investor_irr, fill = suboptimal)) +
  geom_histogram(binwidth = 0.001) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = c("#999999", "#D55E00")) +
  geom_vline(xintercept = investor_irr_ci$investor_irr_05, linetype = "dashed") + 
  geom_vline(xintercept = investor_irr_ci$investor_irr_95, linetype = "dashed") +
  labs(
      title = "Investor IRR Distribution",
      subtitle = str_glue(
        "90% confidence estimate: {round(investor_irr_ci$investor_irr_05*100, 1)}% to ",
        "{round(investor_irr_ci$investor_irr_95*100, 1)}%"
      ),
      x = "Investor IRR",
      y = "",
      caption = "Source: Simulation Results"
    ) + 
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    ) +
    guides(
      fill = FALSE
    )

suboptimal_pct = simulation_data %>% 
  summarize(suboptimal_pct = mean(investor_irr < 0.10)) %>%
  pull(suboptimal_pct)

investor_summary_chart = ivesntor_summary_tbl %>%
  mutate(suboptimal = ifelse(investor_irr_outcome == "less than 10%", "yes", "no")) %>%
  mutate(pct_label = pct %>% scales::percent(accuracy = 1)) %>%
  ggplot(aes(pct, investor_irr_outcome, fill = suboptimal, label = pct_label)) +
  geom_col() + 
  geom_label() +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = c("#999999", "#D55E00")) +
  labs(
      title = "Investor IRR Outcomes",
      subtitle = str_glue(
        "{suboptimal_pct %>% scales::percent(accuracy = 1)} chance of an ",
        "investor IRR of less than 10%"
      ),
      x = "% of Simulation Trials",
      y = ""
    ) + 
    guides(
      fill = FALSE
    )

plots = list(investor_summary_chart, investor_irr_dist)
p = wrap_plots(plots)
  
  p %>% 
    ggsave(
      filename = str_glue("plots/investor_outcomes.png"),
      width = 10, height = 5
    )
```



## Export Data

```{r}
gs4_auth(secrets$stern_email)

investor_irr_ci_tbl %>%
  sheet_write(
        ss = secrets$sheet_id,
        sheet = "Sensitivity Analysis - Part 2 (Investor IIR CI)"
    )

results_tbl %>%
  select(occupancy_distribution, summary) %>%
  unnest(summary) %>%
    sheet_write(
        ss = secrets$sheet_id,
        sheet = "Sensitivity Analysis - Part 2 (Investor IIR Buckets)"
    )
```

